Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase Retrieval with Automatic Differentiation

### 2023-04-14 src/tex/ms.tex
[showyourwork] overleaf sync

In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical optics models to astronomical data. By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as automatic differentiation or `autodiff' \cite{Margossian2018} features that enable high-dimensional optimization and inference with gradient descent or \ac{hmc}. In the present paper, we focus on the particular problem of estimating the \ac{prf} and \ac{psf} simultaneously - i.e. joint flat field calibration and phase retrieval. In Paper~II, we will show how autodiff enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end. And in subsequent papers, we will use this for end-to-end deconvolution of high angular resolution imaging with an unknown \ac{psf}. We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource. This paper has been compiled using ShowYourWork \cite{Luger2021}, so that all figures link to the code used to produce them.

@@

This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm. This total model requires the calculation of 60 individual \ac{psf}s over a $1024\times1024$ detector from a $512\times512$ pixel wavefront. This calculation is vastly dominated by the propagation of the wavefront using a 2-sided MFT, which requires two matrix multiplications for each individual \ac{psf} of sizes $[N_0 \times N_i] \cdot [N_i \times N_i]$ and $[N_0 \times N_i] \cdot [N_i \times N_o]$ where $N_i = 512$ is the input size of the wavefront and $N_o = 1024$ is the output size of the wavefront. This takes a total of $\sim$ 1.25 seconds to evaluate on an Apple M1 CPU, with all operations performed using FP32 precision default to \jax. FP32 was chosen as this problem does not require precision over a large dynamic range, however FP64 calculations can be performed as required. We also apply 4 integer-pixel dithers of +- 20 pixels in the (x,y) directions, equivalent to a 1 $\lambda/D$ dither for a total of 5 images with the total flux spread evenly across each image. This helps the over-parameterisation consequential from having a free parameter for each pixel. Using pixel-integer dithers similarly allows for a single PSF calculation on a slightly larger detector plane (1064x1064 in this case) and then sampling the smaller 1024x1024 regions to produce the full set of 5 images.

>>>>

In this series of papers, we present a new software package, \dlux, for fitting high-dimensional parametrized physical optics models to astronomical data. By using the Python library \textsc{Jax}\cite{jax}, we obtain \ac{gpu} hardware acceleration, as well as automatic differentiation or `autodiff' \cite{Margossian2018} features that enable high-dimensional optimization and inference with gradient descent or \ac{hmc}. In the present paper, we focus on the particular problem of estimating the \ac{prf} and \ac{psf} simultaneously - i.e. joint flat field calibration and phase retrieval. In Paper~II, we will show how autodiff enables improvements to hardware \textit{design}, directly calculating and optimizing the Fisher information and therefore fundamental figures of merit of an optical system end-to-end. And in subsequent papers, we will use this for end-to-end deconvolution of high angular resolution imaging with an unknown \ac{psf}. We make \dlux available as open-source software on GitHub\footnote{\href{https://github.com/LouisDesdoigts/dLux}{github.com/LouisDesdoigts/dLux}}, and encourage interested readers to use and contribute to this package as a community resource. This paper has been compiled using ShowYourWork \cite{Luger2021}, so that all figures link to the code used to produce themcan be found \href{https://github.com/LouisDesdoigts/instrumental_calibration/tree/main}{here}.

@@

This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545\,nm, 595\,nm and 645\,nm. This total model requires the calculation of 60 individual \ac{psf}s over a $1024\times1024$ detector from a $512\times512$ pixel wavefront. This calculation is vastly dominated by the propagation of the wavefront using a 2-sided MFT, which requires two matrix multiplications for each individual \ac{psf} of sizes $[N_o \times N_i] \cdot [N_i \times N_i]$ and $[N_o \times N_i] \cdot [N_i \times N_o]$ where $N_i = 512$ is the input size of the wavefront and $N_o = 1024$ is the output size of the wavefront. This takes a total of $\sim$ 1.25 seconds to evaluate on an Apple M1 CPU, with all operations performed using FP32 precision default to \jax. FP32 was chosen as this problem does not require precision over a large dynamic range, however FP64 calculations can be performed as required. We also apply 4 integer-pixel dithers of +- 20 pixels in the (x,y) directions, equivalent to a 1 $\lambda/D$ dither for a total of 5 images with the total flux spread evenly across each image. This helps the over-parameterisation consequential from having a free parameter for each pixel. Using pixel-integer dithers similarly allows for a single PSF calculation on a slightly larger detector plane (1064x1064 in this case) and then sampling the smaller 1024x1024 regions to produce the full set of 5 images.
